Human Judgement Forecasting of COVID-19 in the UK

### 2023-04-15 manuscript/manuscript.tex
Updates from Overleaf

Infectious disease modelling and forecasting has attracted wide-spread attention during the COVID-19 pandemic and helped inform decision making in public health organisations and governments \cite{cramerEvaluationIndividualEnsemble2021, venkatramananUtilityHumanJudgment2022}. 
% Beginning in March 2020, forecasts for different COVID-19 targets have been systematically collated by Forecast Hubs in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. 
Most forecasts used to inform decision making were based on computational models of COVID-19, but some authors also explored human judgement forecasting as an alternative or in combination \cite{recchiaHowWellDid2021, mcandrewExpertJudgmentModel2022, bosseComparingHumanModelbased2022, mcandrewChimericForecastingCombining2022}.

@@

In some situations, human judgement forecasting may have advantages relative to computational models. Human judgment may be particularly useful to provide timely forecasts in situations where data is sparse and many parameters are unknown or structural breaks have occurred. Humans are generally able to answer a broad set of question (such as for example the likelihood that a given actor will take some specified action) or can take factors into account that are hard to encode in a computational model. On the other hand, human judgement forecasting is difficult to scale due to the time and effort required, and humans may be at a disadvantage at tasks that strongly benefit from the ability to perform complex computations. Also, the use of human judgement forecasts by decision makers may be complicated by the lack of clarity of the basis on which they were made.

@@

This paper represents a follow-up study to \citet{bosseComparingHumanModelbased2022} in the United Kingdom with forecasts made over the course of thirteen weeks between May and August 2021. Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the estimated effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a simulation model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

The median number of unique participants in any given week was 17, the minimum was 6 and the maximum was 51. This was higher than the number of participants in \cite{bosseComparingHumanModelbased2022} (which had a median number of 6, a minimum of 2, and a maximum 10). With respect to the number of submissions from an individual participant, we observed similar patterns as \cite{bosseComparingHumanModelbased2022}: An individual forecaster participated on average in 2.6 weeks out of 13. The median number of submissions from a single individual was one, meaning that similar to \citep{bosseComparingHumanModelbased2022} most forecasters dropped out after their first submission. Only five participants submitted a forecast in ten or more weeks and only two submitted a forecast in all thirteen weeks, one of whom is an author on this study. The number of direct forecasts (median: 13 for cases and 12 for deaths) was higher than the number of $R_t$ forecasts (median: 6 for both cases and deaths) in all weeks (see Figure \ref{fig:num-forecasters}A). The median number of "non-experts" (11 for cases, 10 for deaths) was higher than the median number of "experts" (8 for cases and deaths) (see Figure \ref{fig:num-forecasters}B).

@@

In terms of WIS on the log scale, all human forecasting approaches outperformed the Forecast Hub ensemble for two week ahead forecasts of cases (see Figure \ref{fig:performance}). WIS values relative to the Hub ensemble (=1) were 0.91 for the combined crowd ensemble, 0.96 for the direct crowd forecasts and 0.93 for the $R_t$ forecasts (see Table \ref{tab:scores}). 
In terms of WIS on the natural scale, however, the Hub ensemble was ahead of all human forecasting approaches. Relative WIS values on the natural scale for two week ahead forecasts were 1.05 for the combined crowd ensemble, 1.03 for the direct crowd forecasts and 1.2 for the $R_t$ forecasts. 
This discrepancy between performance on the log scale can be attributed to the fact that case forecasts of the Hub ensemble tended to be lower than forecasts of the human judgement forecasting approaches (see Figure \ref{fig:forecasts-scores}). On the natural scale, this led to smaller overprediction penalties on the natural scale, putting it ahead of human forecasts (see Figure \ref{fig:performance}A,C). On the log scale, however, it led to large penalties for underprediction.

@@

In terms of relative model ranks for two week ahead forecasts, the Hub ensemble and the $R_t$ showed a higher variance than the combined crowd ensemble and the direct forecasts (See Figure \ref{fig:performance-ranks}), despite forecasts being about the same or more dispersed (see Figure \ref{fig:performance}). Both the Hub ensemble and the $R_t$ forecast were more often in first place than other approaches (4 times each, both on the log and on the natural scale). However, they were also most often in the last place (Hub ensemble: 6 on the log scale and 5 on the natural scale, $R_t$: 5 on the log scale and 6 on the natural scale). The direct forecasts placed relatively equally in places 1-4. The crowd ensemble never placed fourth, but also had the lowest number of first places (2, both on the log and the natural scale). Model ranks only changed marginally when switching between the log and the natural scale.

@@

In terms of WIS on the log scale for two week ahead predictions of deaths, the combined crowd ensemble (0.97) and the direct crowd forecasts (0.99) were marginally ahead of the Hub ensemble, while the $R_t$ forecasts performed noticeably worse (1.98) (see Figure \ref{fig:performance}D and Table \ref{tab:scores}). For the Hub ensemble, the dispersion component played by far the largest role, while this was less the case for the human forecasts, which got higher penalties from both over- and underprediction. Interestingly, combining the $R_t$ forecasts and the direct forecasts led to an ensemble that performed better than either of them alone on the log scale. 
In terms of WIS on the natural scale, only the direct forecasts (0.89) performed better for two week ahead death predictions than the Hub ensemble, while the combined crowd ensemble performed a bit worse (1.06) and the $R_t$ forecasts again noticeably worse (2.1).

% Maybe say something about increasing forecast horizons. But the pattern is unclear.

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/comparison-direct-rt-individual.png}
\caption{\bf{Comparison of predictive performance of individual forecasters using either the direct forecasting or $R_t$ interface}. Comparisons are based only on those instances where forecasters have submitted a prediction using both interfaces. The absolute level for a given forecaster relative to others is not meaningful as forecasters differ in the amounts of forecasts they have submitted and when.}
\label{fig:comparison-direct-rt-individual}
\end{figure*}
For cases, where participants could observe the case forecast implied by their $R_t$ forecast, predictive performance was similar between corresponding direct and $R_t$ forecasts for most forecasters who had submitted both (see Figure \ref{fig:comparison-direct-rt-individual}). For deaths, where forecasters could not see the incidence forecast implied by their $R_t$ forecast or manually adjust the case fatality rate, performance of the $R_t$ forecasts was significantly worse. From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). In particular, $R_t$ forecasts for deaths were worse than the corresponding direct death forecasts for most forecasters (see Figure \ref{fig:comparison-direct-rt-individual}). Changing from the direct forecasting method to $R_t$ forecasting for cases tended to improve scores for better forecasters and decrease scores for worse forecasters, although sample sizes are very small.

@@

When comparing "experts" and "non-experts", the ensemble of "experts" was better calibrated for forecasts two weeks ahead (see Figure \ref{fig:performance-experts}. For cases two week ahead, "experts" achieved better scores than "non-experts" both on the log and on the natural scale. WIS values \textit{relative to the combined crowd ensemble} were 1.08 for "experts" and 1.14 for "non-experts" on the log scale and 1.06 for "experts" and 1.07 for "non-experts" on the natural scale (see Table \ref{tab:scores-experts}).  
For two week ahead forecasts of deaths, "experts" performed worse than "non-experts" in terms of WIS on the log scale (WIS relative to the combined crowd ensemble: 1.29 vs. 1.06), but better on the natural scale (1.03 vs. 1.14). Both the "expert"- and the "non-expert"-ensemble had similar proportions of $R_t$ forecasts (mean of 32\% for "experts" and 32.2\% for "non-experts" across cases and deaths together).

@@

% \paragraph{Summary}
In this paper, we presented a follow-up study to \cite{bosseComparingHumanModelbased2022}, analysing human judgement forecasts of cases of and deaths from COVID-19 in the United Kingdom submitted to the European COVID-19 Forecast Hub between the 24th of May and the 16th of August 2021. Human judgement forecasts were generated using two different forecasting approaches, a) direct forecasts of cases and deaths and b) forecasts of the effective reproduction number $R_t$, which were based on estimates from an open source effective reproduction number estimation model and also relied on this model to generate simulated reported cases and deaths and an extension which estimated a fixed case fatality ratio and a log normal delay between cases and reports.

@@

%Calibration
In line with results from previous work \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, we found almost all forecasts for cases to be underdispersed (i.e. too narrow and overconfident). Empirical coverage for death forecasts was higher for all forecasting approaches except the $R_t$ forecasts.

@@

In their original study, conducted in Germany and Poland, \citet{bosseComparingHumanModelbased2022} found that humans outperformed an ensemble of computational models when predicting cases, but not when predicting deaths. They hypothesised that computational models might have an advantage over human forecasters when predicting deaths, benefiting from the ability to model the delays and epidemiological relationships between different leading and lagged indicators. \citet{mcandrewChimericForecastingCombining2022} similarly found in their study that humans performed comparably to an ensemble of computational models for cases, but not for predictions of deaths of COVID-19. 
Results in our study do not directly support this pattern, but given the low number of observations also do not provide strong evidence against it. In this study, the combined crowd ensemble performed better than the Hub ensemble on both cases and deaths on the log scale, and worse on the natural scale. Direct forecasts, which would be most comparable to the forecasts in \citep{bosseComparingHumanModelbased2022}, performed worse than the Hub ensemble on cases and better on deaths. 
Relatively good performance of direct human judgement forecasts for death forecasts compared to previous studies may be related to changes in the age distribution of cases during the study period and the rise of the Delta variant in the UK starting in the beginning of May. The Delta variant had a higher case fatality ratio (CFR) than previous variants \cite{linDiseaseSeverityClinical2021}. One possible hypothesis for the relatively good performance of human forecasts for deaths compared to previous studies might be that some models submitted to the Forecast Hub may have been more negatively affected by the changes in CFR during the study period than human forecasters. A confounding factor, when comparing results from this study and the one in Germany and Poland directly, is that we used a median ensemble to combine individual forecasts here, while the earlier study had used a mean ensemble.

@@

This study explored a novel method of forecasting infectious diseases that combines a human forecast of the estimated effective reproduction number $R_t$ with epidemiological modelling to map the $R_t$ forecast to a forecast of cases and deaths. One appeal of this approach is that the forecaster can directly forecast the generative process. Computational modelling would then take care of dealing with details such as reporting delays, generation intervals, day of the week periodicity, or the relationship between different indicators. This could help reduce cognitive load, and make it easier to synthesise various sources in information into a single forecast, at least for forecasters who have an intuitive understanding of $R_t$. Anecdotally, forecasters familiar to the authors reported high satisfaction with the forecasting experience. In our study, $R_t$ forecasts of cases were comparable to direct forecasts, with a tendency for good forecasts to improve when using the $R_t$ method and worse forecasts to deteriorate even more. Sample sizes, however, were very low. 
Given that forecasters could simulate cases in the app, it is also possible that forecasters were in reality directly forecasting cases. 
$R_t$ forecasts of deaths (which forecasters could not see in the app) were noticeably worse than direct forecasts of deaths. The computational model underlying our $R_t$ forecasts of deaths estimated a constant CFR and delay distribution using the last 4 weeks of data, therefore updating relatively slowly to new circumstances and not evolving at all over the forecast period. The CFR, however, changed quite quickly during the study period, both due to the rise of the Delta variant \citep{bastIncreasedRiskHospitalisation2021}, as well as changes in the age distribution of cases in Summer 2021, in parts related to the European Football Championship \citep{dehningImpactEuro20202023}. Forecasters had no way of inspecting the death forecast implied by their $R_t$ forecast and also had no way to adjust the CFR manually, likely impacting forecast accuracy. Giving human forecasters the ability to adjust the CFR and other model parameters would have increased complexity of the interface, but would have solved issues with the assumptions of the underlying model. Alternatively, a more complex model could have been used which allowed for time-varying CFR estimates and forecast these changes over the forecast horizon. Another important limitation is that we didn't have full sample trajectories of the $R_t$-values predicted by forecasters. Rather, trajectories had to be constructed based on the distributions provided for the different forecast horizons, which likely affected forecasts. One potential way to disentangle the effect of the convolution model from the forecasts would have been to use the human forecasts for cases as an input to the computational model, which could then have estimated deaths as a convolution of cases over a delay distribution. Future work could expose forecasters to different such options with the aim of separating effects of the user interface from ones related to the structure of the underlying computational model.

@@

Past expert forecasts of COVID-19 \cite{recchiaHowWellDid2021} had found predictions from experts to outperform those of non-experts. In our study, an ensemble of self-reported "experts" outperformed an ensemble of "non-experts" when forecasting cases two weeks ahead, both on the log scale and on the natural scale. When forecasting deaths two weeks ahead, "experts" performed worse than "non-experts" on the log scale, but better on the natural scale. Forecasts for "experts" tended to be better calibrated than non-experts.  However results should be taken with care in light of relatively low sample sizes (median of 11 "non-experts" for cases and 10 for deaths, median of 8 "experts" for cases and deaths), and given that expert status was self-reported. Furthermore, we only asked for professional involvement in a field related to infectious disease modelling, not specifically for familiarity with modelling of COVID-19 in the UK, and only offered participants a binary choice.

%results should be interpreted with caution. Overall, it seems that ensembles comprised of either human forecasts or computational ensembles are able to produce forecasts of similar quality.

% \paragraph{Susceptibility of results}

Overall, results of our study should be taken with caution due to a number of important limitations. Firstly, our study was restricted to one location and to a relatively short period of thirteen weeks. Secondly, there were many confounding factors that likely influence results. These include for example the fact that different participants made forecasts at different points in time (with the median forecaster only submitting a single forecast) and that subgroups of interest (e.g. "experts", or $R_t$ forecasts) had different numbers of forecasters. 
In most instances, differences in scores between forecast approaches were small compared to the variance of scores within a single approach. In addition, there were many researcher degrees of freedom that could influence findings, for example how individual forecasts were combined to an ensemble. Results were strongly influenced by the choices made during the evaluation and for example changed depending on which horizon was chosen for comparisons and whether or not forecasts were transformed before scoring. 
Prizes to the human forecasters, for example were paid out based on the combined WIS on the log scale across all horizons and forecast targets. Had we chosen to instead measure WIS on the natural scale, rankings and payouts would have been different.

@@

The results of our study seem broadly consistent with previous studies on human judgement forecasting of COVID-19 and suggest that human crowd ensembles and an ensemble of computational models are able to produce forecasts of similar quality. 
One possible interpretation is that a mixed crowd of human forecaster can produce a viable alternative or addition to an ensemble of mathematical models created by experts. Another interpretation is that an ensemble of automated models can produce forecasts over the course of several years that are on par with that of an engaged crowd of human forecasters. All previous studies comparing human judgement forecasts and computational models only ran over short periods of time and the majority of them struggled with recruitment and upkeep. Meanwhile, COVID-19 Forecast Hubs have attracted continuous submissions for almost three years and were able to consistently provide forecasts of comparable quality.

Our findings do not suggest that humans are necessarily at a general disadvantage compared to computational models at predicting death numbers, but evidence in both directions is limited and this is made particularly complex as our study took place during a period of time when CFR estimates were changing rapidly. Despite evaluations being public, it remains a challenge to properly incentivise contributors to Forecast Hubs to regularly update their forecasting methodology in order to maximise utility, predictive performance, or both. Combining human judgement and epidemiological modelling by mapping $R_t$ forecasts to case and death numbers has not yielded competitive forecasts for deaths in this study. However, we only presented a prototype of a forecasting approach, which, while having appealing properties, proved challenging to implement. Subsequent iterations and improvements could likely achieve better results. More research is required to obtain a better understanding of the role of subject matter expertise in infectious disease forecasting. Our results underline that it is difficult to evaluate forecast performance devoid of context that helps inform what a good or a bad forecast is. Different ways to look at the data let different forecasts appear good or bad. Forecast evaluation therefore either needs to be clearly informed by the needs of forecast consumers to determine what a good forecast is, or it needs a broad array of perspectives to provide a wholistic picture. Furthermore, evaluating forecasts post-hoc leaves the researchers with many degrees of freedom to make decisions that strongly affect which models look good and there is a high risk of allowing for motivated reasoning. More emphasis should be put on measures that prevent this, e.g. by establishing common standards for evaluations, pre-registering studies, and making it a norm to display a variety of metrics.

>>>>

Infectious disease modelling and forecasting has attracted wide-spread attention during the COVID-19 pandemic and helped inform decision making in public health organisations and governments \cite{cramerEvaluationIndividualEnsemble2021, venkatramananUtilityHumanJudgment2022}.  
Most forecasts used to inform decision making were based on computational models of COVID-19, but some authors also explored human judgement forecasting as an alternative or in combination \cite{recchiaHowWellDid2021, mcandrewExpertJudgmentModel2022, bosseComparingHumanModelbased2022, mcandrewChimericForecastingCombining2022}.

@@

In some situations, human judgement forecasting may have advantages relative to computational models. Human judgment may be particularly useful to provide timely forecasts in situations where data is sparse and many parameters are hard to quantify. Humans are also generally able to answer a broad set of question (such as for example the likelihood that a given actor will take some specified action) and can take factors into account that are hard to encode in a computational model. On the other hand, human judgement forecasting is difficult to scale due to the time and effort required, and humans may be at a disadvantage at tasks that strongly benefit from the ability to perform complex computations. Also, the use of human judgement forecasts by decision makers may be complicated by the lack of clarity of the basis on which they were made.

@@

This paper represents a follow-up study to \citet{bosseComparingHumanModelbased2022} in the United Kingdom with forecasts made over the course of thirteen weeks between May and August 2021. Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2022}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country, in a different time period, and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the estimated effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a simulation model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

The median number of unique participants in any given week was 17, the minimum was 6 and the maximum was 51. This was higher than the number of participants in \cite{bosseComparingHumanModelbased2022} (which had a median number of 6, a minimum of 2, and a maximum 10). With respect to the number of submissions from an individual participant, we observed similar patterns as \cite{bosseComparingHumanModelbased2022}: An individual forecaster participated on average in 2.6 weeks out of 13. The median number of submissions from a single individual was one, meaning that similar to \citep{bosseComparingHumanModelbased2022} most forecasters dropped out after their first submission. Only five participants submitted a forecast in ten or more weeks and only two submitted a forecast in all thirteen weeks, one of whom is an author on this study (S. Abbott). The number of direct forecasts (median: 13 for cases and 12 for deaths) was higher than the number of $R_t$ forecasts (median: 6 for both cases and deaths) in all weeks (see Figure \ref{fig:num-forecasters}A). The median number of "non-experts" (11 for cases, 10 for deaths) was higher than the median number of "experts" (8 for cases and deaths) (see Figure \ref{fig:num-forecasters}B).

@@

In terms of WIS on the log scale, all human forecasting approaches outperformed the Forecast Hub ensemble for two week ahead forecasts of cases (see Figure \ref{fig:performance}). WIS values relative to the Hub ensemble (=1) were 0.91 for the combined crowd ensemble, 0.96 for the direct crowd forecasts and 0.93 for the $R_t$ forecasts (see Table \ref{tab:scores}). In contrast, in terms of WIS on the natural scale the Hub ensemble outperformed all human forecasting approaches. Relative WIS values on the natural scale for two week ahead forecasts were 1.05 for the combined crowd ensemble, 1.03 for the direct crowd forecasts and 1.2 for the $R_t$ forecasts. 
The discrepancy between performance on the log and natural scale  can be attributed to case forecasts from the Hub ensemble tending to be lower than forecasts from human judgement approaches (see Figure \ref{fig:forecasts-scores}). On the natural scale, this resulted in smaller overprediction penalties, putting it ahead of human forecasts (see Figure \ref{fig:performance}A,C). On the log scale, however, it led to large penalties for underprediction.

@@

In terms of relative model ranks for two week ahead forecasts, the Hub ensemble and the $R_t$ forecast showed a higher variance than the combined crowd ensemble and the direct forecasts (See Figure \ref{fig:performance-ranks}), despite forecasts being about the same or more dispersed (see Figure \ref{fig:performance}). Both the Hub ensemble and the $R_t$ forecast were more often in first place than other approaches (4 times each, both on the log and on the natural scale). However, they were also most often in the last place (Hub ensemble: 6 on the log scale and 5 on the natural scale, $R_t$: 5 on the log scale and 6 on the natural scale). The direct forecasts placed relatively equally in places 1-4. The crowd ensemble never placed fourth, but also had the lowest number of first places (2, both on the log and the natural scale). Aggregated model ranks only changed marginally when switching between the log and the natural scale (see Figure \ref{fig:performance-ranks}.

@@

In terms of WIS on the log scale for two week ahead predictions of deaths, the combined crowd ensemble (0.97) and the direct crowd forecasts (0.99) were marginally ahead of the Hub ensemble, while the $R_t$ forecasts performed noticeably worse (1.98) (see Figure \ref{fig:performance}D and Table \ref{tab:scores}). For the Hub ensemble, the dispersion component played by far the largest role, while this was less the case for the human forecasts, which got higher penalties from both over- and underprediction. Combining the $R_t$ forecasts and the direct forecasts led to an ensemble that performed better than either of them alone on the log scale despite the poor overall performance of the $R_t$ forecasts. 
In terms of WIS on the natural scale, only the direct forecasts (0.89) performed better for two week ahead death predictions than the Hub ensemble, while the combined crowd ensemble performed slightly worse (1.06) and the $R_t$ forecasts again noticeably worse (2.1).

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/comparison-direct-rt-individual.png}
\caption{\bf{Comparison of predictive performance of individual forecasters using either the direct forecasting or $R_t$ interface}. Comparisons are based only on those instances where forecasters have submitted a prediction using both interfaces. The absolute level for a given forecaster relative to others is not meaningful as forecasters differ in the amounts of forecasts they have submitted and when.}
\label{fig:comparison-direct-rt-individual}
\end{figure*}
For cases, where participants could observe the case forecast implied by their $R_t$ forecast, predictive performance was similar between corresponding direct and $R_t$ forecasts for most forecasters who had submitted both (see Figure \ref{fig:comparison-direct-rt-individual}). For deaths, where forecasters could not see the incidence forecast implied by their $R_t$ forecast or manually adjust the case fatality rate, performance of the $R_t$ forecasts was significantly worse. From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). In particular, $R_t$ forecasts for deaths were worse than the corresponding direct death forecasts for most forecasters (see Figure \ref{fig:comparison-direct-rt-individual}). Changing from the direct forecasting method to $R_t$ forecasting for cases tended to improve scores for better forecasters and decrease scores for worse forecasters, although sample sizes and the size of the observed effect are both small.

@@

When comparing two week ahead forecasts from "experts" and "non-experts", the ensemble of "experts" was better calibrated (see Figure \ref{fig:performance-experts}. For cases, "experts" achieved better scores than "non-experts" both on the log and on the natural scale. WIS values \textit{relative to the combined crowd ensemble} were 1.08 for "experts" and 1.14 for "non-experts" on the log scale and 1.06 for "experts" and 1.07 for "non-experts" on the natural scale (see Table \ref{tab:scores-experts}).  
For deaths, "experts" performed worse than "non-experts" in terms of WIS on the log scale (WIS relative to the combined crowd ensemble: 1.29 vs. 1.06), but better on the natural scale (1.03 vs. 1.14). Both the "expert"- and the "non-expert"-ensemble had similar proportions of $R_t$ forecasts (mean of 32\% for "experts" and 32.2\% for "non-experts" across cases and deaths together).

@@

% \paragraph{Summary}
In this paper, we presented a follow-up study to \cite{bosseComparingHumanModelbased2022}, analysing human judgement forecasts of cases of and deaths from COVID-19 in the United Kingdom submitted to the European COVID-19 Forecast Hub between the 24th of May and the 16th of August 2021. Human judgement forecasts were generated using two different forecasting approaches, a) direct forecasts of cases and deaths and b) forecasts of the effective reproduction number $R_t$, which were based on estimates from an open source effective reproduction number estimation model and also relied on this model, along with a second model relating cases and deaths from the same source, to simulate reported cases and deaths.

@@

%Calibration
In line with results from previous work \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, we found almost all forecasts for cases to be underdispersed (i.e. too narrow/overconfident). Empirical coverage for death forecasts was higher than the corresponding coverage for cases for all forecasting approaches except the $R_t$ forecasts.

@@

In their original study, conducted in Germany and Poland, \citet{bosseComparingHumanModelbased2022} found that humans outperformed an ensemble of computational models when predicting cases, but not when predicting deaths. They hypothesised that computational models might have an advantage over human forecasters when predicting deaths, benefiting from the ability to model the delays and epidemiological relationships between different leading and lagged indicators. \citet{mcandrewChimericForecastingCombining2022} similarly found in their study that humans performed comparably to an ensemble of computational models for cases, but not for predictions of deaths of COVID-19. 
Results in our study do not directly support this pattern, but given the low number of observations also do not provide strong evidence against it. In this study, the combined crowd ensemble performed better than the Hub ensemble on both cases and deaths on the log scale, and worse on the natural scale. Direct forecasts, which would be most comparable to the forecasts in \citep{bosseComparingHumanModelbased2022}, performed worse than the Hub ensemble on cases and better on deaths. 
During the study period, the case fatality ratio (CFR) changed quite quickly. This was related to the rise of the Delta variant in the UK in the beginning of May with its higher case fatality rate \citep{bastIncreasedRiskHospitalisation2021}. In addition, the age distribution of cases changed over the course of the study period in Summer 2021, in parts related to the European Football Championship \citep{dehningImpactEuro20202023}.
One possible hypothesis for the relatively good performance of human forecasts for deaths compared to previous studies might be that some models submitted to the Forecast Hub may have been more negatively affected by the changes in CFR during the study period than human forecasters or have been slower to update . A confounding factor, when comparing results from this study and the one in Germany and Poland directly, is that we used a median ensemble to combine individual forecasts here, while the earlier study used a mean ensemble.

@@

This study explored a novel method of forecasting infectious diseases that combines a human forecast of the estimated effective reproduction number $R_t$ with epidemiological modelling to map the $R_t$ forecast to a forecast of cases and deaths. One appeal of this approach is that the forecaster can directly forecast the generative process and how they believe it is affected by interventions and changes in behaviour. Computational modelling then takes care of dealing with details such as reporting delays, generation intervals, day of the week periodicity, and the relationship between different indicators. This could help reduce cognitive load, and make it easier to synthesise various sources in information into a single forecast, at least for forecasters who have an intuitive understanding of $R_t$. Though we note all of these modelling steps and the construction of the model itself requires the human constructing the model to make assumptions. Anecdotally, forecasters familiar to the authors reported high satisfaction with the forecasting experience. In our study, $R_t$ forecasts of cases were comparable to direct forecasts, with a tendency for good forecasts to improve when using the $R_t$ method and worse forecasts to deteriorate even more. Sample sizes, however, were very low. 
Given that forecasters could simulate cases in the app, it is also possible that forecasters were in reality directly forecasting cases. 
$R_t$ forecasts of deaths (which forecasters could not see in the app) were noticeably worse than direct forecasts of deaths. The computational model underlying our $R_t$ forecasts of deaths estimated a constant CFR and delay distribution using the last 4 weeks of data, therefore updating relatively slowly to new circumstances and was assumed to be constant over the four week forecast horizon. However, as mentioned before, the CFR evolved during the study period. Forecasters had no way of inspecting the death forecast implied by their $R_t$ forecast and also had no way to adjust the CFR manually, likely impacting forecast accuracy. Giving human forecasters the ability to adjust the CFR and other model parameters would have increased complexity of the interface, but would have solved issues with the assumptions of the underlying model. Alternatively, a more complex model could have been used which allowed for time-varying CFR estimates and forecast these changes over the forecast horizon though this approach may still have struggled to cope with the rapid changes observed during the study period. Another important limitation is that we didn't have full sample trajectories of the $R_t$-values predicted by forecasters. Rather, trajectories had to be constructed based on the distributions provided for the different forecast horizons, which likely negatively affected forecasts. One potential way to disentangle the effect of the convolution model from the $R_t$ forecasts would have been to use the human forecasts for cases as an input to the second computational model, which could then have simulated deaths. Future work could expose forecasters to different combinations of these options with the aim of separating effects of the user interface from ones related to the structure of the underlying computational model.

@@

Past expert forecasts of COVID-19 \cite{recchiaHowWellDid2021} had found predictions from experts to outperform those of non-experts. In our study, an ensemble of self-reported "experts" outperformed an ensemble of "non-experts" when forecasting cases two weeks ahead, both on the log scale and on the natural scale. When forecasting deaths two weeks ahead, "experts" performed worse than "non-experts" on the log scale, but better on the natural scale. Forecasts for "experts" tended to be better calibrated than non-experts.  However results should be taken with care in light of relatively low sample sizes (median of 11 "non-experts" for cases and 10 for deaths, median of 8 "experts" for cases and deaths), and given that expert status was self-reported. Furthermore, we only asked for professional involvement in a field related to infectious disease modelling, not specifically for familiarity with modelling of COVID-19 in the UK, and only offered participants a binary choice. However, as we used ad-hoc recruitment in our networks many of these self-identified experts are likely to be infectious disease modellers.

% \paragraph{Susceptibility of results}
Overall, results of our study should be taken with caution due to a number of important limitations. Firstly, our study was restricted to one location and to a relatively short period of thirteen weeks. Secondly, there were many confounding factors that likely influence results. These include the fact that different participants made forecasts at different points in time (with the median forecaster only submitting a single forecast) and that subgroups of interest (e.g. "experts", or $R_t$ forecasts) had different numbers of forecasters. 
In most instances, differences in scores between forecast approaches were small compared to the variance of scores within a single approach. In addition, there were many researcher degrees of freedom that could influence findings, for example how individual forecasts were combined to create an ensemble. Results were influenced by choices made during the evaluation with, for example, some conclusions  depending on forecast horizon and the transformation used prior to scoring. 
Highlighting this, prizes to the human forecasters were paid out based on the combined WIS on the log scale across all horizons and forecast targets. Had we chosen to instead measure WIS on the natural scale, or to forecast only cases and continue to score on the log scale, rankings and payouts would have been different.

@@

The results of our study are broadly consistent with previous studies on human judgement forecasting of COVID-19 and suggest that human crowd ensembles and an ensemble of computational models are able to produce forecasts of similar quality. 
One interpretation of these findings is that a mixed crowd of human forecaster can produce a viable alternative or complement to an ensemble of mathematical models created by experts. An altnerative interpretation is that an ensemble of automated models can produce forecasts over the course of several years that are on par with that of an engaged crowd of human forecasters. This study, and all previous studies, comparing human judgement forecasts and computational models only ran over short periods of time and the majority of them struggled with recruitment and upkeep. Meanwhile, COVID-19 Forecast Hubs have attracted continuous submissions for almost three years and were able to consistently provide forecasts of comparable quality.

Our findings do not suggest that humans are necessarily at a general disadvantage compared to computational models at predicting reported deaths, but evidence in both directions is limited and this is made particularly complex as our study took place during a period of time when CFR estimates were changing rapidly. Despite evaluations being public, it remains a challenge to properly incentivise contributors to Forecast Hubs to regularly update their forecasting methodology in order to maximise utility, predictive performance, or both. Combining human judgement and epidemiological modelling by mapping $R_t$ forecasts to case and death numbers has not yielded competitive forecasts for deaths in this study. However, we only presented a prototype of a forecasting approach, which, while having appealing properties, proved challenging to implement. Subsequent iterations and improvements could likely achieve better results. More research is required to obtain a better understanding of the role of subject matter expertise in infectious disease forecasting. Our results underline that it is difficult to evaluate forecast performance devoid of context that helps inform what a good or a bad forecast is. Different ways to look at the data let different forecasts appear better or worse. Forecast evaluation therefore either needs to be clearly informed by the needs of forecast consumers to determine what a good forecast is, or it needs a broad array of perspectives to provide a wholistic picture as we have attempted to present in this work. Furthermore, evaluating forecasts post-hoc leaves the researchers with many degrees of freedom to make decisions that affect which models look good and there is a risk of allowing for motivated reasoning. More emphasis should be put on measures that prevent this, e.g. by establishing common standards for evaluations, pre-registering studies, and making it a norm to display a variety of standard metrics.
